### 2018-02-08 results.tex
added nodes-depth-attribute figures and worked on results section

As explained in the methodology, the first step is to perform clustering process on the dataset in order to properly label the samples according to the classes to be used in the decision tree algorithm. Figure \ref{fig:clusters} presents a sample of the results obtained after the constrained \kmeans{} clustering process is completed, including subspacing. Recall that there are a total of 550 subspaces. Fig.~\ref{fig:clusters} shows three examples for each one of the types of subspaces, that is, three for each one of the 2-, 3-, and 4-feature analyses. However, since it is not practical or possible to present three- or higher-dimensional plots, we display the results from a two-dimensional perspective into the the subspaces by picking 2 metrics at a time. (Note that in the only cases in which this is a direct view into the clustering results is in the cases of the 2-feature analysis.)

% There are some characteristics of the results shown in Fig.~\ref{fig:clusters} worth

subspaces for an ordered sequence of metric pairs from C1 to C11 (see Table \ref{tab:metrics}). Although it would be impractical to discuss all the possible combinations, some points can be made from those shown in this figure. For instance, not all feature combinations lead to clearly distinguishable clusters. In this sample we include well-partitioned sets such as C3-C4, C5-C6, C7-C8, and some others where the data-points are mixed across classes, such as subspaces C1-C2 and C10-C11. Note also that although some data distribution may hint the existence of clusters, they may nonetheless be irrelevant because knowledge about one feature does not add insight about the other. That is the case of the C9-C10 subspace. Conversely, there are subspaces with a clear correlation between the data, but with indistinguishable clusters. That is the case of the C1-C2 subspace. These latter two types of subspaces do not contribute information to the clustering process \citep{Dy_2004_MLR}.

\begin{figure*}[ht!]
	\centering
	\includegraphics[width=\textwidth]{figures/pdf/figure-05}
	\caption{A sample of the results from the multi-dimensional clustering analysis for pairs of consecutive metrics C1 through C11. These results are part of the 55 possible sub-spaces with 2 features (metrics). The cluster centers are shown with a cross symbol. Although for some pairs it may be obvious, the clusters 1 through 4 do not necessarily correspond in that order to the validation classes poor, fair, good, and excellent. The color version of this figure is available only in the electronic edition.}
	\label{fig:clusters}
\end{figure*}

Figure \ref{fig:boxed-clusters} shows the statistical distribution of the data in the form of box-plots for each metric and component of motion once partitioned into the four validation classes. Separately, we have prepared similar plots for the different velocity models used in the simulations from where we obtained our dataset. As in Fig.~\ref{fig:data-box-plot} before the clustering, and in Fig.~\ref{fig:data-box-plot} after the clustering process, we observe that the velocity models and the components of motion do not have a strong influence on the statistical distribution of the data, that is why we merge all the data samples into a single dataset at the before processing the data and we continue forward with that approach after clustering and labeling.

In total, the initial dataset is partitioned into groups of 816, 1253, 879 and 76 data samples for the poor, fair, good and excellent classes, respectively. These groups are shown in Fig.~\ref{fig:count-classes}, which also show the contribution of the component of motion to the final count of samples in each class for illustrative purposes. The main aspect to notice in this figure and the quantities just provided is the fact that the number of samples in the excellent class is significantly less than those of the poor, fair, and good classes.

@@

As a consequence, before moving on with the decision tree analysis, it is necessary to do resampling on the subset of the excellent class. In order to maintain a good number of samples for the other three classes, we used the oversampling approach described in the previous section. To that end, we replicated data randomly by a $\times$10 factor until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes}. By common standards, an oversample rate of $\times$10 is considered acceptable \citep{Weiss_2003_JAIR}. (Arguably, we could have as well undersampled the fair class, but we deemed that unnecessary.) Once this process was completed we went on with the decision tree analysis.

\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{figures/pdf/figure-07}
	\caption{Number of data samples in each class (poor, far, good, excellent) after conducting a multi-dimensional constrained \kmeans{} clustering process using subspace analysis for 2, 3, and 4 features (GOF metrics). Here, for illustrative purposes, the filling of the bars distinguishes the component of motion of the compared signals as classified. The dashed-line bar indicates the number of samples in the excellent class after oversampling.}
	\label{fig:count-classes}
\end{figure}

In total we generated 20,000 trees using the C5.0 algorithm for combinations of the parameters CF and $S_{\min}$. CF was chosen to vary between 0 and 1 at intervals of size 0.01, and $S_{\min}$ between 1 and 200 at unitary intervals. Not surprisingly, for such a small size intervals, the algorithm often reached the same tree for different CF and $S_{\min}$ values. Therefore, in reality, from the 20,000 we ran the algorithm, we only found 66 unique trees. For each one of these unique threes, we computed the effectiveness factor $F_1$ from equation (\ref{eq:f}) and extracted the total number of nodes in the trees and their depth.

Figure... (working)

>>>>

As explained in the methodology, the first step is to carry out a clustering process on the dataset in order to properly label the samples according to the classes to be used in the decision tree algorithm. Figure \ref{fig:clusters} presents a sample of the results obtained after applying the constrained \kmeans{} method, including subspacing. Recall that there are a total of 550 subspaces. Figure \ref{fig:clusters} shows three examples for each one of the subspaces considered, that is, three for each one of the 2-, 3-, and 4-feature subspace analyses. However, since it is not practical or possible to present three- or higher-dimensional plots, we display the results from a two-dimensional perspective into the the subspaces by picking 2 metrics at a time. (Note that the only cases in which this is a direct view into the clustering results are the 2-feature analysis cases.)

Some aspects of Fig.~\ref{fig:clusters} are worth discussing. Although in most plots all four clusters are clearly distinguishable, there are others in which that is not the case. In the C5-C9 and C7-C9 combinations in the 3-feature analysis, for instance, the excellent cluster has a limited presence. This is due to the influence of the lower C9 values (see Fig.~\ref{fig:data-box-plot}) on the correlations with C5 and C7. This is not necessarily a bad thing, as it helps to understand the different relationships between the various GOF metrics. In the best of cases, the combinations C1-C2 and C5-C8 in the 2- and 4-feature analysis, for instance, exhibit an almost perfect proportionality between the metrics. These are called redundant relationships, where knowledge about a feature provides a direct view into another feature's results. On the other hand, the combination C1-C7 in the 2-feature analysis presents an example of an irrelevant relationship, where knowledge about a feature provides no insight about the other. Subspaces with redundant relationships are useful for reducing the number of features. Irrelevant relationships, however, do not contribute much to the whole process \citep{Dy_2004_MLR}.

\begin{figure*}[ht!]
	\centering
	\includegraphics[width=\textwidth]{figures/pdf/figure-05}
	\caption{A sample of the results from the multi-dimensional clustering analysis showing the clusters from a two-dimensional perspective into the relationships between different GOF metrics. In each case, the labels near the upper-left corner indicate the features considered in the subspace analysis, and the features associated to the horizontal ($x$) and vertical ($y$) axes in the plot. The top, middle, and bottom rows correspond to 2-, 3-, and 4-feature subspace analyses. In each case, the poor, fair, good, and excellent clusters are indicated with circle, triangle, square, and cross symbols, respectively. The color version of this figure is available only in the electronic edition.}
	\label{fig:clusters}
\end{figure*}

Figure \ref{fig:boxed-clusters} shows the statistical distribution (box-plots) of the samples once the dataset is partitioned into the four validation classes. This is similar to Fig.~\ref{fig:data-box-plot}, but after the clustering process is completed. Separately, we prepared similar plots to look at the influence of the velocity models and components of motions on the results of the clustering process and, as observed before, they had no significant differences with the aggregate of all the samples in the dataset.

@@

In total, the clustering process results in 816, 1253, 879 and 76 data samples for the poor, fair, good and excellent classes, respectively. These groups are shown in Fig.~\ref{fig:count-classes}. As it can be seen in this figure, the number of samples in the excellent class is significantly less than those in the poor, fair, and good classes. Therefore, before moving on with the decision tree analysis, it is necessary to resample the subset of the excellent class. We used the oversampling approach described in the previous section. We replicated data from the excellent class randomly by a $\times$10 factor until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes} with the dashed line. By common standards, an oversample rate of $\times$10 is considered acceptable \citep{Weiss_2003_JAIR}. (Arguably, we could have as well undersampled the fair class, but we deemed that unnecessary.) Once this process was completed we went on with the decision tree analysis.

\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{figures/pdf/figure-07}
	\caption{Number of data samples in each class (poor, far, good, excellent) after conducting a multi-dimensional constrained \kmeans{} clustering process using subspace analysis for 2, 3, and 4 features (GOF metrics). Here, for illustrative purposes, the filling of the bars distinguishes the component of motion of the compared signals as classified. The dashed-line bar indicates the number of samples in the excellent class after oversampling. The color version of this figure is available only in the electronic edition.}
	\label{fig:count-classes}
\end{figure}

In total we generated 20,000 trees using the C5.0 algorithm for all possible combinations of the parameters CF and $S_{\min}$, where CF was chosen to vary between 0 and 1 at intervals of size 0.01, and $S_{\min}$ was chosen to vary between 1 and 200 at unitary intervals. Despite our choice for small intervals, the algorithm often reached recurrent tree structures for different CF and $S_{\min}$ values. Therefore, in reality, from the 20,000 combinations for which we ran the algorithm, only 66 unique trees were found.

For each one of these unique trees, we computed the effectiveness factor $F_1$ from equation (\ref{eq:f}), and extracted the total number of nodes in the trees and their depth. Figure \ref{fig:nodes-depths} shows the results of $F_1$ for all the trees and its distribution in terms of the number of attributes used in the trees as a function of the number of nodes and the depth of each tree. Recall that we are interested in finding a sequence of decisions (represented by disjunctive decision nodes in a tree) that can lead to good GOF predictions (i.e, high values of $F_1$) using a reduced number of attributes (GOF metrics). In general, all the trees obtained with the C5.0 algorithm are good in terms of the effectiveness factor ($F_1$ values close to 1). Then our choice comes down to using a reduced number of metrics. Having several trees with 2, 3, and 4 metrics (as opposed to 11), the following factor in the decision is choosing trees with algorithms using a small number of steps to reach the prediction. This is given by a combination between the depth of the tree and the number of nodes in the tree (i.e., trees with low complexity).

\begin{figure*}[t]
	\centering
	\includegraphics[width=\textwidth]{figures/pdf/figure-08}
	\caption{Trees effectiveness results in terms of the factor ($F_1$) from equation \ref{eq:f} as indicated by colored dots distributed by the number of attributes (GOF metrics) as a function of \textbf{(a)} the number of nodes, and \textbf{(b)} the depth of the trees. The rings around the left-most bottom dots indicate selected trees shown in Fig.~\ref{fig:trees}. The color version of this figure is available only in the electronic edition.}
	\label{fig:nodes-depth}
\end{figure*}

Based on these criteria, we selected three candidate trees: T1, T2, and T3. These trees are shown in Fig.~\ref{fig:trees}.

\begin{figure*}
	\centering
	% \includegraphics[width=\textwidth]{figures/pdf/figure-09}
	\caption{}
	\label{fig:trees}
\end{figure*}
